The resurgence of the linear optics \sihui{quantum} interferometer --- recent advances \& applications

### 2018-01-12 paper.tex
Added sparsity concepts to Section 6.

In the last decade, single photons have been produced mainly through a nonlinear optical process know as spontaneous parametric down-conversion (SPDC), a nonlinear optical process. SPDC produces single photons in correlated pairs that has been a staple source of entanglement. Through developments in production, the number of high quality indistinguishable photons that can be simultaneously produced have increased steadily over the years. Single-photon pairs, once a novelty, are now run-of-the-mill business for quantum optical labs, and three or four photons have become commonplace \cite{bib:Spagnolo12, bib:Tillmann13, bib:Spring13,bib:Broome13}. Experiments using up to ten photons have been demonstrated \cite{bib:WangChen16, bib:Chen17}. We have also seen the meteoric rise of quantum dots in the role for producing single photons \cite{bib:Ding16,bib:WangHe16, bib:213601}. They are a scalable source of single photons which are highly indistinguishable, and are produced on demand--a significant advantage over SPDC which is an inherently probabilistic process.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\columnwidth]{configMZI}
\caption{ \sihui{(a) Efficient decomposition of arbitrary linear optics networks into a sequence of beamsplitters and phase-shifters represented as subunitaries, $D_i$. These subunitaries are controlled via their photon amplitudes ($\alpha_{i,j}$) and phases ($\phi_{i,j}$). This setup describes a generalized multi-port Mach-Zehnder interferometer, and is a universal linear-optical processing unit. (b) A state-of-art implementation of a universal linear-optical processing unit is shown. The processor has 6 inputs entering from the left, and 6 outputs leaving to the right, and is constructed from a cascade of 15 Mach-Zehnder interferometers. The classical control and processing unit sets the parameters for the interferometer, and actively cools it. Adapted from \cite{bib:Carolan15}.}} \label{fig:Carolan}	
\end{figure}

Alternatively, Mach-Zedner interferometers can also be employed as building blocks instead of beamsplitters and phase-shifters \cite{bib:Englert2001}. \sihui{Such a decomposition and its implementation on an integrated photonic chip are shown in Fig.~\ref{fig:Carolan}. } Later, it was shown that any nontrivial beamsplitter, that does more than permuting modes or adding phases to them, is universal for linear optics \cite{bib:Bouland2014}. However, they do not provide an explicit construction for arbitrary unitaries.

@@

State tomography is achieved by measuring expectation values in a complete basis for the state Hilbert space. For one qubit, for example, a complete basis is the Pauli basis, which wields the completeness relation,}
\begin{align}
\hat\rho = \text{tr}({\hat\rho}) + \text{tr}(\hat{X}\hat{\rho})\hat{X} + \text{tr}({\hat{Y}\hat\rho})\hat{Y} + \text{tr}(\hat{Z}{\hat\rho})\hat{Z}.
\end{align}
\peter{These traces correspond exactly to the expectation values in the respective bases, which are easily determined using separable measurements. Process tomography employs state tomography as a primitive, applying an equivalent completeness relation, but across the space of quantum processes, which may be decomposed into a basis input to output states. For a single qubit process this may be represented as,}
\begin{align} \label{eq:process_matrix}
\mathcal{E}(\hat\rho) = \sum_{i,j=1}^4 \chi_{i,j} \hat{\sigma}_i\hat\rho\,\hat{\sigma}_j^\dag.
\end{align}
\peter{which generalises to an arbitrary number of qubits.Here $\chi$ is the process matrix in the Pauli basis.}
%\cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}.

>>>>

In the last decade, single photons have been produced mainly through \sihui{nonlinear optical processes know as spontaneous parametric down-conversion (SPDC) and spontaneous four-wave mixing (SFWM) \cite{bib:Eisaman11,bib:Xiong16,bib:Solntsev17}.} \sihui{Both SPDC and SFWM produce} single photons in correlated pairs that has been a staple source of entanglement. \sihui{However, SPDC uses up one pump photons per pair, while SFWM requires two.} Through developments in production, the number of high quality indistinguishable photons that can be simultaneously produced have increased steadily over the years. Single-photon pairs, once a novelty, are now run-of-the-mill business for quantum optical labs, and three or four photons have become commonplace \cite{bib:Spagnolo12, bib:Tillmann13, bib:Spring13,bib:Broome13}. Experiments using up to ten photons have been demonstrated \cite{bib:WangChen16, bib:Chen17}. We have also seen the meteoric rise of quantum dots in the role for producing single photons \cite{bib:Ding16,bib:WangHe16, bib:213601}. They are a scalable source of single photons which are highly indistinguishable, and are produced on demand--a significant advantage over SPDC which is an inherently probabilistic process.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\columnwidth]{configMZI}
\caption{ \sihui{(a) A decomposition of arbitrary linear optics networks into a sequence of subunitaries, $D_i$, made up of Mach-Zehnder interferometers, and is a universal linear-optical processing unit. These subunitaries are controlled via their photon amplitudes ($\alpha_{i,j}$) and phases ($\phi_{i,j}$). (b) A state-of-art implementation of a universal linear-optical processing unit is shown. The processor has 6 inputs entering from the left, and 6 outputs leaving to the right, and is constructed from a cascade of 15 Mach-Zehnder interferometers. The classical control and processing unit sets the parameters for the interferometer, and actively cools it. Adapted from \cite{bib:Carolan15}.}} \label{fig:Carolan}	
\end{figure}

Alternatively, \sihui{Mach-Zehnder} interferometers can also be employed as building blocks instead of beamsplitters and phase-shifters \cite{bib:Englert2001}. \sihui{Such a decomposition and its implementation on an integrated photonic chip are shown in Fig.~\ref{fig:Carolan}. } Later, it was shown that any nontrivial beamsplitter, that does more than permuting modes or adding phases to them, is universal for linear optics \cite{bib:Bouland2014}. However, they do not provide an explicit construction for arbitrary unitaries.

@@

State tomography is achieved by measuring expectation values in a complete basis for the state Hilbert space. For one qubit, for example, a complete basis is the Pauli basis, which wields the completeness relation,}
\begin{align}
\hat\rho = \text{tr}({\hat\rho}) + \text{tr}(\hat{X}\hat{\rho})\hat{X} + \text{tr}({\hat{Y}\hat\rho})\hat{Y} + \text{tr}(\hat{Z}{\hat\rho})\hat{Z}.
\end{align}
\peter{These traces correspond exactly to the expectation values in the respective bases, which are easily determined using separable measurements. \sihui{Sometimes, it is possible to be able to do a tomography from an incomplete number of measurements, {\it i.e.} a number that is less than the number of expectation values, and it is possible using compressed sensing provided some sparsity conditions are met \cite{bib:GLFBE2010,bib:Titchener16,bib:Oren16}.} Process tomography employs state tomography as a primitive, applying an equivalent completeness relation, but across the space of quantum processes, which may be decomposed into a basis input to output states. For a single qubit process this may be represented as,}
\begin{align} \label{eq:process_matrix}
\mathcal{E}(\hat\rho) = \sum_{i,j=1}^4 \chi_{i,j} \hat{\sigma}_i\hat\rho\,\hat{\sigma}_j^\dag.
\end{align}
\peter{which generalises to an arbitrary number of qubits. Here $\chi$ is the process matrix in the Pauli basis.}
%\cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}.
